A UK Case Study on\\Cybersecurity Education and Accreditation

### 2019-06-23 FIE2019-final.tex
Loads of changes and tweaks

\section{Cybersecurity: for all or for specialists?}

@@

\section{Research Questions}
There are various levels of specialism at which cybersecurity education can be addressed.
\begin{enumerate}
\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry? Minor point but maybe worth mentioning e.g. being safe online, accessing public services, etc}})
%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design. BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
\item The general computer science graduate.
\item The generalist computer science masters graduate.
\item The specialist computer science graduate.
\item The specialist computer science masters graduate.
\end{enumerate}
The focus of this paper is on 2-3: the general computer science graduate.
\begin{description}
\item[RQ1]What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?
\item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?
\item[RQ3]Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?
\end{description}

\subsection{Research Approach}
A UK-focused case study-based approach is adopted in this project. As is common in case study-based research, many alternative case studies could have been chosen. The cases are illustrative rather than comprehensive in terms of the available case studies or challenges. The cases were evaluated to articulate the progress made and highlight opportunities for future developments in education and practice.

@@

In order to address RQ1 and RQ2 a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated. Namely, an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.

Finally, in order to address RQ3 cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK). As discussed in the current situation section, in this jurisdiction mandating cybersecurity within PSRB accreditation is at a reasonably mature stage.

\subsection{Current State of Play}
The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to RQ2:
\begin{quote}
{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
given the manner in which the topics are pervasive throughout other KAs}}
\end{quote}
It proposes 9 ``core'' hours and 63.5 distributed across the other KAs.
%\pagebreak - to fix URL issues
\par
Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, indicates
\begin{quote}
{\emph{Universities suffer shortcomings, as well. Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity. There is a big catch, however. Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}
\end{quote}

But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: the paper quotes that 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping. As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.

It is at least plausible to attribute this difference to differences in the accreditation regimes, as the external circumstances, governmental pressures, and learned society curricula are all similar.
\begin{description}
\item[UK]BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle). More precisely, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
%BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body. ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots''. This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
%\cite[chapter 7]{Davenportminor2019a}.  % need a better reference
\end{description}

\section{Challenges: RQ1 and RQ2}

%What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
%{\bf JHD to research ACM general curriculum;}
The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states 3 Tier-1 and 6 Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.

The UK's official body of knowledge resource, the CyBOK project\cite{Bristol2019a}, has produced reference documentation for some (5  final, 7 for comment out of a planned total of 19)   % JHD updated figures 21/6/2019
knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).

@@

\subsection{Industry standards - Payments Card Industry Data Security Standards (PCI DSS)}\label{sec:PCIDSS}
Is teaching cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}). \cite{Weiss:2013:THC:2527148.2527180} suggests that there are benefits from teaching this discipline area in a more practical manner.  There are a number of ways a more practical manner could be action, one would be by the inclusion of pertinent cybersecurity standards in the curricula.

The PCI DSS \cite{PCI2018b} is one such standard. PCI DSS underpins all processing of credit/debit cards. Nevertheless, they are very rarely mentioned in generalist computer scientist courses. This would not matter so much if everyone handling payments data were sent by their employers on an effective PCI DSS course. However, the payments business is now so spread across websites, often run by small and medium enterprises (SME), or non-specialists.  Even larger enterprises are not immune: \cite{Barth2018a} reports that the recent British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance.

@@

\subsection{Educational Resources}\label{sec:EDResource}
\subsubsection{Structured Query Language (SQL) Injection}\label{sec:SQL}
It is 15 years since \cite{Guimaraesetal2004} 
wrote ``All the topics listed above should be presented in the first
Database Course'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}. SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon (\url{https://xkcd.com/327/}, dating back to 2007 according to the Internet Archive) and website (\url{http://bobby-tables.com/}). Nevertheless SQL injection is still a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.  \cite[the UK's definitive reference]{Bristol2018a} states ``a wide range of attack techniques for exploiting SQL
injection or script injection are known and documented.''.

Clearly such a major weakness should be well-taught.  In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recommended textbooks.

% Drop2019 was awarded the (unique) best paper at SIGCSE2019 - not sure what to make of this.
Hence \cite{Drop2019} analysed the database textbooks used by  44 of the top 50 computer science
departments in the USA. % (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed). 
There were seven such books, but three books accounted for the 36 of the 44 universities. Five of the seven (30 of the 44) had no mention of SQL injection. On the other two, the more popular one has a seriously flawed discussion\footnote{``However, the paper implies that using parameters is equivalent to using a function to add escape characters
around user input. This is incorrect, as using parameters allows
SQL statements to be pre-compiled, and prevents any user input
from being interpreted as code, while escaping user input is not
recommended as a sole defense since imperfect escape functions
can easily be subverted.'' \cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``However, the fact that the first
example should not be used is not discussed until two pages after
the example in the text, and is not mentioned at all in the caption or
on the page where the figure appears. This means a student who is
skimming the text looking for an example to modify for their own
code could simply copy the code that first appears in the example,
without being aware that this is in fact an example of what they
shouldn't do.'' \cite{Drop2019}}.
\par
This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself.\footnote{Fixing this is a suggestion for future work.}
\subsubsection{The case of Java}\label{sec:Java}
A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for  Java textbooks. Indeed, many such books go nowhere near security applications.  But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources. \cite{Mengetal2018a} analysed 503 cybersecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.  53\% were about the Spring Security framework (\url{https://projects.spring.io/spring-security/}), dominated by authentication (45\%). The discussion \cite[\S4.3.1]{Mengetal2018a} of cross-site request forgery (CSRF) is especially worrying.  By default, Spring implicitly enables protection against this. But all the accepted answers to CSRF-related failures simply suggested disabling the check. There were no negative comments about this, and indeed a typical response is ``Adding \verb!csrf().disable()!
solved the issue!!! I have no idea why it was enabled by default''. As of writing (16 January 2018) there were no negative comments about this disabling of a vital security feature.
\par
This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}). Their first finding was
\begin{quote}
644 out of the 1,429 inspected answer posts
(45\%) are insecure, meaning that insecure suggestions
popularly exist on SO. Insecure answers dominate, in
particular, the SSL/TLS category [355 insecure versus 150 secure, i.e. $>70$\%].
\end{quote}

\subsubsection{Android}\label{sec:Android}
A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for Android textbooks. \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. The key finding was this.
\begin{quote}
We found that 15.4\% of all 1.3 million Android applications
contained security-related code snippets from
Stack Overflow. Out of these 97.9\% contain at least one
insecure code snippet.
\end{quote}
Two caveats (in opposite directions) should be noted. The labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure. On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
\par
The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}. Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``Improve documentation: Clarify what you can(not) copy/paste''. 
\subsubsection{Agile}
A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for ``Agile'' textbooks. Many authors have found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
\par
Agile's preference for functionality over non-functional requirements is clearly displayed in practice. \cite{Naiakshinaetal2017a} asked 20 student developers to imagine they
were part of a team working on creating a social networking site for
our university and to implement a password storage mechanism for this. 10 (``primed'') were explicitly told that the storage had to be secure and 10 (``unprimed'') were not. None of the unprimed ones implemented any security.
\subsubsection{Informal Resources}\label{sec:informal}
The web abounds with informal resources: tutorials and code snippets. How good are these, and how good are people at using these? This has been looked at by \cite{Unruhetal2017a}, This research took the top five results from Google for six queries. Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses. Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection. Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students' (and indeed others') use of such Informal Resources.

\subsection{Staff}
It is well known that cybersecurity skills are in short supply, for example \cite{Page2018a}.
\begin{quote}
% JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years. In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills. This had climbed to 51\% at the beginning of this year. Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.
\end{quote}
The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
%JHD possibly to write about his recruitment experiences.  ? more IoC input?

In the United Kingdom there is a prominence of job adverts to recruit academic staff with specialisms in cybersecurity. The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations. As universities expand their cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.  A recent example had a professor of cybersecurity, two senior academic positions and 2 junior academic positions in one advert. There are examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year. There are also examples of  cybersecurity subject groups moving en masse from one university to another.

%Not sure if we want to raise something about funding of posts ??
%Poaching of staff ??
%Is this too contentious

%{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
\section{Accreditation by PSRBs: RQ3}

In the UK as in most jurisdictions, higher education provision addresses general computer science and specialist cybersecurity courses. A significant number of undergraduate and postgraduate programmes are available in both the areas of computer science and cybersecurity (and closely related fields computer security, digital/computer forensics, etc). In the UK, the Universities and Colleges Admissions Service (UCAS) lists over 40 higher education institutions (HEIs) providing undergraduate qualifications related to cybersecurity for entry in September 2019. An even larger number of HEIs provide study opportunities related to more general computer science. UCAS lists 246 provides for undergraduate programmes related to computer science.

@@

\textbf{1. Not-for-Profit Organisations}
Tech Partnership Degrees is a not-for-profit organisation that provides endorsements to higher education programmes with specific curricula elements aimed at job market requirements. One of the curricula elements is related to cybersecurity. Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business. Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and five Software Engineering for Business Programmes. As such, Tech Partnership Degrees currently have limited impact upon more general computer science education and none upon specialist cybersecurity education.

@@

\textbf{2. National Cyber Security Centre (NCSC)}
The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK. The NCSC publishes and accredits to a number of cybersecurity standards \cite{NCSC2018a}. These standards are linked to the ACM recommendations for curricula \cite{ACM2013a}. To date the major focus of NCSC accreditation has been upon Masters degrees specialising in cybersecurity. More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and computer science degrees with a significant cybersecurity focus \cite{NCSC2018b}.

@@

\textbf{3. Professional Bodies}
The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of computer science and the more specialist area of cybersecurity discipline areas. The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}. These memoranda support the internationalising of the curriculum and promote consistency and parity in computer science education globally.   These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}. Both the IET and The BCS have a long history of expecting coverage of environmental factors within the programmes they accredit. The BCS has for a number of years been expecting significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}. Clearly cybersecurity has been and continues to be part of these expectations.

>>>>

\section{Cybersecurity: For All, Or For Specialists?}

@@

\section{Research Questions}

There are various levels of specialism at which cybersecurity education can be addressed:
\begin{enumerate}[label=(\roman*)]
\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry? Minor point but maybe worth mentioning e.g. being safe online, accessing public services, etc}})
%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design. BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
\item The generalist computer science graduate.
\item The generalist computer science masters graduate.
\item The specialist computer science graduate.
\item The specialist computer science masters graduate.
\end{enumerate}
The focus of this paper is on (ii)--(iii): the general computer science graduate.

This paper thus focuses on three research questions:

\begin{description}
\item[RQ1] {\emph{What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?}}
\item[RQ2] {\emph{Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?}}
\item[RQ3] {\emph{Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?}}
\end{description}

\subsection{Research Approach}

A UK-focused case study-based approach is adopted in this project. As is common in case study-based research, many alternative case studies could have been chosen. The cases are illustrative rather than comprehensive in terms of the available case studies or challenges. The cases were evaluated to articulate the progress made and highlight opportunities for future developments in education and practice.

@@

In order to address {\emph{RQ1}} and {\emph{RQ2}} a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated. Namely, an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.

Finally, in order to address {\emph{RQ3}} cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK). As discussed in the current situation section, in this jurisdiction mandating cybersecurity within PSRB accreditation is at a reasonably mature stage.

\subsection{Current State of Play}

The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to RQ2:
\begin{quote}
{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
given the manner in which the topics are pervasive throughout other KAs}}
\end{quote}
It proposes 9 ``core'' hours and 63.5 distributed across the other KAs.
%\pagebreak - to fix URL issues
\par
Nevertheless, the situation on the ground in the USA is different~\cite{Ackerman2019a}:
\begin{quote}
{\emph{Universities suffer shortcomings, as well. Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity. There is a big catch, however. Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}
\end{quote}

The UK situation is distinctly different: 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping~\cite[Table 1]{Ruiz2019a}. As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.

It is at least plausible to attribute this difference to differences in the accreditation regimes, as the external circumstances, governmental pressures, and professional body/learned society curricula are all similar.

\begin{description}
\item[UK]BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle). More precisely, accredited degrees have been expected to demonstrate coverage of ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information systems}}'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
%BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body. ABET is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''. This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
%\cite[chapter 7]{Davenportminor2019a}.  % need a better reference
\end{description}

\section{Challenges: {\emph{RQ1}} and {\emph{RQ2}}}

%What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
%{\bf JHD to research ACM general curriculum;}
The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states three Tier-1 and six Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.

The UK's official body of knowledge resource, the CyBOK project\cite{Bristol2019a}, has produced reference documentation for some (as of June 2019: five final, seven for comment, out of a planned total of 19)   % JHD updated figures 21/6/2019
knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).

@@

\subsection{Industry Standards: PCI DSS}\label{sec:PCIDSS}

Is teaching cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}). \cite{Weiss:2013:THC:2527148.2527180} suggests that there are benefits from teaching this discipline area in a more practical manner.  There are a number of ways a more practical manner could be action, one would be by the inclusion of pertinent cybersecurity standards in the curricula.

The Payments Card Industry Data Security Standards (PCI DSS)~\cite{PCI2018b} is one such standard. PCI DSS underpins all processing of credit/debit cards. Nevertheless, they are very rarely mentioned in generalist computer scientist courses. This would not matter so much if everyone handling payments data were sent by their employers on an effective PCI DSS course. However, the payments business is now so spread across websites, often run by small and medium enterprises (SME), or non-specialists.  Even larger enterprises are not immune: \cite{Barth2018a} reports that the recent British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance.

@@

\subsection{Educational Resources}\label{sec:EDResource}

\subsubsection{SQL Injection}\label{sec:SQL}

It is 15 years since \cite{Guimaraesetal2004} 
wrote ``All the topics listed above should be presented in the first
Database Course'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}. SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon (\url{https://xkcd.com/327/}, dating back to 2007 according to the Internet Archive) and website (\url{http://bobby-tables.com/}). Nevertheless SQL injection is still a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.  \cite[the UK's definitive reference]{Bristol2018a} states ``a wide range of attack techniques for exploiting SQL
injection or script injection are known and documented.''.

Clearly such a major weakness should be well-taught.  In general it is hard to determine what is taught, but a reasonable proxy for this is the content of recommended textbooks.

% Drop2019 was awarded the (unique) best paper at SIGCSE2019 - not sure what to make of this.
Hence \cite{Drop2019} analysed the database textbooks used by  44 of the top 50 computer science
departments in the USA. % (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed). 
There were seven such books, but three books accounted for the 36 of the 44 universities. Five of the seven (30 of the 44) had no mention of SQL injection. On the other two, the more popular one has a seriously flawed discussion\footnote{``{\emph{However, the paper implies that using parameters is equivalent to using a function to add escape characters
around user input. This is incorrect, as using parameters allows
SQL statements to be pre-compiled, and prevents any user input
from being interpreted as code, while escaping user input is not
recommended as a sole defense since imperfect escape functions
can easily be subverted.}}''~\cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``{\emph{However, the fact that the first
example should not be used is not discussed until two pages after
the example in the text, and is not mentioned at all in the caption or
on the page where the figure appears. This means a student who is
skimming the text looking for an example to modify for their own
code could simply copy the code that first appears in the example,
without being aware that this is in fact an example of what they
shouldn't do.}}''~\cite{Drop2019}}.
\par
This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself.\footnote{Fixing this is a suggestion for future work.}

\subsubsection{The Case of Java}\label{sec:Java}

A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for  Java textbooks. Indeed, many such books go nowhere near security applications.  But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources. \cite{Mengetal2018a} analysed 503 cybersecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.  53\% were about the Spring Security framework (\url{https://projects.spring.io/spring-security/}), dominated by authentication (45\%). The discussion \cite[\S4.3.1]{Mengetal2018a} of cross-site request forgery (CSRF) is especially worrying.  By default, Spring implicitly enables protection against this. But all the accepted answers to CSRF-related failures simply suggested disabling the check. There were no negative comments about this, and indeed a typical response is ``Adding \verb!csrf().disable()!
solved the issue!!! I have no idea why it was enabled by default''. As of writing (16 January 2018) there were no negative comments about this disabling of a vital security feature.
\par
This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}). Their first finding was:

\begin{quote}
{\emph{644 out of the 1,429 inspected answer posts
(45\%) are insecure, meaning that insecure suggestions
popularly exist on SO. Insecure answers dominate, in
particular, the SSL/TLS category}} [355 insecure versus 150 secure, i.e. $>70$\%].
\end{quote}

\subsubsection{Android}\label{sec:Android}

A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for Android textbooks; \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. The key finding was this:
\begin{quote}
{\emph{We found that 15.4\% of all 1.3 million Android applications
contained security-related code snippets from
Stack Overflow. Out of these 97.9\% contain at least one
insecure code snippet.}}
\end{quote}

Two caveats (in opposite directions) should be noted. The labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure. On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
\par
The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}. Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``{\emph{Improve documentation: Clarify what you can(not) copy/paste}}''.

\subsubsection{Agile}

A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for ``Agile'' textbooks. Many authors have found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
\par
Agile's preference for functionality over non-functional requirements is clearly displayed in practice. \cite{Naiakshinaetal2017a} asked 20 student developers to imagine they
were part of a team working on creating a social networking site for
our university and to implement a password storage mechanism for this. 10 (``primed'') were explicitly told that the storage had to be secure and 10 (``unprimed'') were not. None of the unprimed ones implemented any security.

\subsubsection{Informal Resources}\label{sec:informal}

The web abounds with informal resources: tutorials and code snippets. How good are these, and how good are people at using these? This has been looked at by \cite{Unruhetal2017a}, This research took the top five results from Google for six queries. Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses. Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection. Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students' (and indeed others') use of such informal resources.

\subsection{Staff}

It is well known that cybersecurity skills are in short supply~\cite{schneider2013}. For example:
\begin{quote}
% JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
{\emph{Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years. In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills. This had climbed to 51\% at the beginning of this year. Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.}}~\cite{Page2018a}
\end{quote}
The ESG survey is international, but ESG have confirmed that the UK figures are very similar.

In the UK, there is a prominence of job adverts to recruit academic staff with specialisms in cybersecurity. The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations. As universities expand their cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time. Recent example have included a professor of cybersecurity, two senior academic positions and two junior academic positions in one advert. There are other examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year; there are also examples of cybersecurity subject groups moving en masse from one university to another.

%Not sure if we want to raise something about funding of posts? Poaching of staff ??
%Is this too contentious
%TC: probably out of scope for this paper

%{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
\section{Accreditation by PSRBs: {\emph{RQ3}}}

In the UK as in most jurisdictions, higher education provision addresses general computer science and specialist cybersecurity courses. A significant number of undergraduate and postgraduate programmes are available in both the areas of computer science and cybersecurity (and closely related fields computer security, digital/computer forensics, etc). In the UK, the Universities and Colleges Admissions Service (UCAS) lists over 40 higher education institutions providing undergraduate qualifications related to cybersecurity for entry in September 2019. An even larger number of institutions provide study opportunities related to more general computer science. UCAS lists 246 provides for undergraduate programmes related to computer science.

@@

\textbf{1. Not-for-Profit Organisations}

Tech Partnership Degrees is a not-for-profit organisation that provides endorsements to higher education programmes with specific curricula elements aimed at job market requirements. One of the curricula elements is related to cybersecurity. Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business. Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and five Software Engineering for Business Programmes. As such, Tech Partnership Degrees currently have limited impact upon more general computer science education and none upon specialist cybersecurity education.

@@

\textbf{2. National Cyber Security Centre (NCSC)}

The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK. The NCSC publishes and accredits to a number of cybersecurity standards \cite{NCSC2018a}. These standards are linked to the ACM recommendations for curricula \cite{ACM2013a}. To date the major focus of NCSC accreditation has been upon Masters degrees specialising in cybersecurity. More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and computer science degrees with a significant cybersecurity focus \cite{NCSC2018b}.

@@

\textbf{3. Professional Bodies/Learned Societies}

BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of computer science and the more specialist area of cybersecurity discipline areas. The accreditation provided by these institutes are underpinned by international initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and the Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}. These memoranda support the internationalising of the curriculum and promote consistency and parity in computer science education globally.   These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}. Both the BCS and the IET have a long history of expecting coverage of environmental factors within the programmes they accredit. The BCS has for a number of years been expecting significant coverage of what it terms  {\emph{Legal, Ethical, Social and Professional Issues}}~\cite{Brooke2018}. Clearly cybersecurity has been and continues to be part of these expectations.
